---
title: "Fit delay from infection to death"
output: html_notebook
---


```{r}
library(tidyverse)
library(cowplot)
library(rstan)
knitr::opts_chunk$set(echo = FALSE)
```


## Import data from Hong Kong

```{r}
check_dates <- function(dates){
  dates[!grepl(x = dates, pattern = '\\d\\d?/\\d\\d?/\\d\\d')]
}
is_valid_date <- function(dates){
  grepl(x = dates, pattern = '\\d\\d?/\\d\\d?/\\d\\d')
}
clean_dates_onset <- function(chdate){
  data_frame(chdate = chdate) %>%
    separate(chdate, into = c('month', 'day', 'year'), sep = '/') %>%
    mutate(date = lubridate::as_date(sprintf('20%02i-%02i-%02i', as.numeric(year), as.numeric(month), as.numeric(day)))) %>%
    pull(date)
}
clean_dates_hosp <- function(chdate){
  data_frame(chdate = chdate) %>%
    separate(chdate, into = c('day', 'month', 'year'), sep = '/') %>%
    mutate(date = lubridate::as_date(sprintf('%04i-%02i-%02i', as.numeric(year), as.numeric(month), as.numeric(day)))) %>%
    pull(date)
}

## Delay from onset to hospitalization from HK data
hkdat <- read_csv('../HK_data/data_24July.csv',
                              col_types = "iiffffff") %>%
  setNames(c('id', 'age', 'sex', 'onset', 'hospital', 'isolation', 'discharge', 'death')) %>%
  select(onset, hospital, age) %>%
  filter(is_valid_date(onset) & is_valid_date(hospital)) %>%
  mutate(onset = clean_dates_onset(onset),
         hospital = clean_dates_hosp(hospital),
         delay = hospital-onset)
```

## Plot by delay
```{r}
hkdat %>%
  ggplot() +
  geom_histogram(aes(x = delay), binwidth = 1) +
  ggtitle('HK delays onset to hospital')
```
## Plot by age
```{r}
hkdat %>%
  mutate(age_bin = cut(age, breaks = seq(0, 100, by = 10), include.lowest = T)) %>%
  ggplot()+
  geom_violin(aes(x = age_bin, y = delay, fill = age_bin), draw_quantiles = c(.1, .5, .9))+
  ylim(c(0,20)) +
  ggtitle('Delay by age') -> vio

vio

vio +
    geom_point(aes(x = jitter(as.numeric(age_bin)), y = jitter(as.numeric(delay))), alpha = .5)
```

## Plot age distribution
```{r}
hkdat %>% 
  group_by(age) %>%
  summarise(n = n()) %>%
  ggplot()+
  geom_bar(aes(x = age, y = n), stat = 'identity')+
  ggtitle('HK hospital admissions with symptom onset date')
```



## Drop negative delays for fitting.
```{r}
  get_shape_rate <- function(mean, vv){
    rate = mean/vv
    shape = mean*rate
    # stopifnot(shape/rate == mean)
    # stopifnot(shape/rate^2 == vv)
    c(shape, rate)
  }

## Empirical delay from onset to hosp
del_onset_to_hosp_hk <- hkdat %>% filter(delay >=0) %>% pull(delay)
```

## Also compare to reported delay from onset to hospitalization from Nishiura et al.
--> Focus on living 
--> Longer for deceased (mean of 24 days)
```{r}
del_onset_to_hosp_Nishiura <- c(
                                rgamma(n = length(del_onset_to_hosp_hk)*.1, get_shape_rate(mean = 3.3, vv = 4.2^2)[1], get_shape_rate(3.3, 4.2^2)[2])) # Deceased (10%)

## Generate synthetic incubation times
del_inf_to_onset <- rgamma(n = length(del_onset_to_hosp_hk), shape = 5.807, scale = 0.948) # From Lauer et al. 2020

delay_vec_hk <- as.numeric(del_inf_to_onset + del_onset_to_hosp_hk)
delay_vec_ni <- as.numeric(del_inf_to_onset + del_onset_to_hosp_Nishiura)
```
## Fit gamma distribution
```{r}
sprintf('mean delay from infection to hospital in hk is %.2f d', mean(delay_vec_hk))
sprintf('mean delay from infection to hospital based on Linton et al is %.2f d', mean(delay_vec_ni))
```


###
```{r}
fit_stan <- function(
  mod = 'fit_gamma.stan',
  del_vec, # vector of delay for each individual in the data
  chains = 1, # Number of MCMC chains to run
  cores = parallel::detectCores(logical = FALSE),
  ...
) {
  library(rstan)
  
  model <- stan_model(mod)
  
  N <- length(del_vec)
  model_input_data <- list(
    N = N,
    x = del_vec
  )
  
  fit <- sampling(
    model, model_input_data, chains = chains, cores = cores,
    ...
  )
}
```


```{r}
fit_stan(mod = '../../code/fit_lognormal.stan', 
               del_vec = delay_vec_hk, 
               chains = 4) -> fit_lnormal_hk
```

```{r}
fit_stan(mod = '../../code/fit_lognormal.stan', 
               del_vec = delay_vec_ni, 
               chains = 4) -> fit_lnormal_ni
```


## Extract lognormal fits
```{r}
params_lognormal_hk <- rstan::extract(fit_lnormal_hk)
params_lognormal_ni <- rstan::extract(fit_lnormal_ni)
```

## Hong Kong Results
```{r}
pairs(fit_lnormal_hk, pars = c("mu", "sigma", "lp__"), las = 1)
print(fit_lnormal_hk, pars = c('mu', 'sigma'))
```
## Linton et al. 2020 results
(living)

```{r}
pairs(fit_lnormal_ni, pars = c("mu", "sigma", "lp__"), las = 1)
print(fit_lnormal_ni, pars = c('mu', 'sigma'))
```

## Plot against data
```{r}
wait_df <- data.frame(wait = floor(delay_vec_ni)) %>%
  group_by(wait) %>%
  summarise(count = n()) %>%
   mutate(kind = 'ni')


post_ni <- mapply(FUN = function(mm, ss) dlnorm(x = wait_df$wait, mm, ss)*length(delay_vec_ni), 
                         mm = params_lognormal_ni$mu,
                         ss = params_lognormal_ni$sigma) %>%
  apply(MARGIN = 1, FUN = function(xx) data.frame(lower = quantile(xx, .025),
                                                  med = quantile(xx, .5),
                                                  upper = quantile(xx, .975))) %>%
  bind_rows() %>%
  mutate(wait = wait_df$wait)



bind_rows(list(
  lognormal = post_ni
), .id = 'dist') %>%
  merge(wait, by = 'wait') %>%
  ggplot()+
  geom_bar(data = wait, aes(x = wait, y = count), stat = 'identity', alpha = .5)+
  geom_line(aes(x = wait, y = med, color = dist), lwd = 1) +
  geom_ribbon(aes(x = wait, ymin = lower, ymax = upper, fill = dist), alpha = .5)+
  ggtitle('Linton (living): Delay from infection to hospitalization', subtitle = sprintf('6.1d incubation; %.2fd onset to hosp\nmean=%.2f, var=%.2f\nlognormal mu = %.2f, sigma=%.2f',
          mean(del_onset_to_hosp_Nishiura), mean(delay_vec_ni), var(delay_vec_ni), mean(params_lognormal_ni$mu), mean(params_lognormal_ni$sigma)))
```


```{r}
wait <- data.frame(wait = floor(delay_vec_hk)) %>%
  group_by(wait) %>%
  summarise(count = n()) %>%
   mutate(kind = 'hk')


post_hk <- mapply(FUN = function(mm, ss) dlnorm(x = wait_df$wait, mm, ss)*length(delay_vec_hk), 
                         mm = params_lognormal_hk$mu,
                         ss = params_lognormal_hk$sigma) %>%
  apply(MARGIN = 1, FUN = function(xx) data.frame(lower = quantile(xx, .025),
                                                  med = quantile(xx, .5),
                                                  upper = quantile(xx, .975))) %>%
  bind_rows() %>%
  mutate(wait = wait_df$wait)



bind_rows(list(
  lognormal = post_hk
), .id = 'dist') %>%
  merge(wait, by = 'wait') %>%
  ggplot()+
  geom_bar(data = wait, aes(x = wait, y = count), stat = 'identity', alpha = .5)+
  geom_line(aes(x = wait, y = med, color = dist), lwd = 1) +
  geom_ribbon(aes(x = wait, ymin = lower, ymax = upper, fill = dist), alpha = .5)+
  ggtitle('HK: delay from infection to hospitalization', subtitle = sprintf('6.1d incubation; %.2fd onset to hosp\nmean=%.2f, var=%.2f\nlognormal mu = %.2f, sigma=%.2f',
          mean(del_onset_to_hosp_hk), mean(delay_vec_hk), var(delay_vec_hk), mean(params_lognormal_hk$mu), mean(params_lognormal_hk$sigma)))
```

## Save the posterior
```{r}
write_rds(params_lognormal_hk, 'delay_infection_to_hosp_posterior.rds')
```

