---
title: "Test sample size"
output: html_notebook
---

Katie Gostic

```{r asis = T}
cat(sprintf('\n## Last run %s', Sys.Date()))
```
```{r}
## Preamble
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
```



## Set observation delays
```{r}
rdelay <- function(nn){
  rlnorm(nn, meanlog = 1.52, sdlog = sqrt(0.39)) + # Incubation period from table 2 of Linton et al.
  runif(nn, 1, 2) %>% ## 1-4d reporting delay
  ceiling()
}

```


## Visualize synthetic data
```{r echo = F}
source('test_sample_size.R')
source('../code/util.R')
parlist <- read_rds('true_pars.rds')
sim_df <- get_obs_ts(rdelay)
get_sim_df('stochastic') %>%
    filter(time < 300) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_line(data = sim_df, aes(x = time, y = incidence), color = 'deepskyblue')+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Epidemic curve') -> inc

get_sim_df('stochastic') %>% 
  filter(time < 300) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_line(data = sim_df, aes(x = time, y = true_rt), color = 'deepskyblue') +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste('Underlying ', R[t], ' values'))) -> R0

cowplot::plot_grid(R0, inc, align = 'hv', nrow = 2)


filter(sim_df, time <200 & time > 175)
```

## Get the times at which Rt crosses one
```{r}
get_rt_cross_1_dates(sim_df)
rt.rise.time = get_rt_cross_1_dates(sim_df) %>% slice(2) %>% pull(time)
```
Debug - Test Cori method.
```{r}
# source('../code/cori.R')
# test_cori <-  merge(
#   get_sim_df('stochastic'), 
#     get_cori(df.in = get_sim_df('stochastic'),
#          icol_name = 'incidence', 
#          out_name = 'test', 
#          window = 1, 
#          GI_mean = parlist$true_mean_GI, 
#          GI_var = parlist$true_var_GI, 
#          wend = F),
#   by = 'time', 
#   all = T)
# 
# test_cori %>%
#   ggplot(aes(x = time))+
#   geom_line(aes(y = true_rt))+
#   geom_line(aes(y = test.mean), color = 'blue')
```



## Test various sample sizes

2000/d - total observed in IL as of 25 Sept.
500/d - total observed in Chicago as of 25 Sept
100/d
50/d
10/d


* Truncate data so the last time of observation falls at the time Rt first rises back above 1, 7, 14 and 21 days later.
* Estimate Rt using each truncated data set.
* Plot results

```{r error=FALSE, warning=FALSE}
# truncate.times <- 250 # rt.rise.time + c(0, 7, 14, 21)
# med_daily_obs <- c(2000, 500, 100, 50, 10)
# test_list = expand.grid(tt = truncate.times, med_cases = med_daily_obs)
# wrapper <- function(tt, med_cases){test_sample_size_end_date(p_obs = 1, 
#                                                   median_cases_per_day = med_cases, 
#                                                   nboot = 100, 
#                                                   last_obs_time = tt, 
#                                                   rdelay = rdelay,
#                                                   min_window = 7)}
# test1 <- mapply(FUN = wrapper, tt = test_list$tt, med_cases = test_list$med_cases, SIMPLIFY = 'list') %>% 
#   apply(2, bind_cols) %>% 
#   bind_rows() %>%
#   mutate(last_day = factor(last_day, levels = sort(truncate.times, decreasing = F)),
#          median_cases_per_day = factor(median_cases_per_day, levels = sort(med_daily_obs, decreasing = F)))
# write_rds(test1, path = 'power_results_7dminWindow.rds')
```


## Repeat without 7d min window
```{r error=FALSE, warning=FALSE}
# truncate.times <- 250 # rt.rise.time + c(0, 7, 14, 21)
# med_daily_obs <- c(2000, 500, 100, 50, 10)
# test_list = expand.grid(tt = truncate.times, med_cases = med_daily_obs)
# wrapper <- function(tt, med_cases){test_sample_size_end_date(p_obs = 1, 
#                                                   median_cases_per_day = med_cases, 
#                                                   nboot = 100, 
#                                                   last_obs_time = tt, 
#                                                   rdelay = rdelay,
#                                                   min_window = 1)}
# test2 <- mapply(FUN = wrapper, tt = test_list$tt, med_cases = test_list$med_cases, SIMPLIFY = 'list') %>% 
#   apply(2, bind_cols) %>% 
#   bind_rows() %>%
#   mutate(last_day = factor(last_day, levels = sort(truncate.times, decreasing = F)),
#          median_cases_per_day = factor(median_cases_per_day, levels = sort(med_daily_obs, decreasing = F)))
# #test2 %>% group_by(median_cases_per_day) %>% summarise(mt = max(time))
# write_rds(test2, path = 'power_results_1dminWindow.rds')
```

## Tune up the smoothing window
```{r, error=F, warning=F}
truncate.times <- 250 # rt.rise.time + c(0, 7, 14, 21)
med_daily_obs <- c(2000, 500, 100, 50, 10)
test_list = expand.grid(tt = truncate.times, med_cases = med_daily_obs)
wrapper <- function(tt, med_cases){test_sample_size_end_date(p_obs = 1, 
                                                  median_cases_per_day = med_cases, 
                                                  nboot = 100, 
                                                  last_obs_time = tt, 
                                                  rdelay = rdelay,
                                                  min_window = 2,
                                                  w.tune = 100)}
test2 <- mapply(FUN = wrapper, tt = test_list$tt, med_cases = test_list$med_cases, SIMPLIFY = 'list') %>% 
  apply(2, bind_cols) %>% 
  bind_rows() %>%
  mutate(last_day = factor(last_day, levels = sort(truncate.times, decreasing = F)),
         median_cases_per_day = factor(median_cases_per_day, levels = sort(med_daily_obs, decreasing = F)))
#test2 %>% group_by(median_cases_per_day) %>% summarise(mt = max(time))
write_rds(test2, path = 'power_results_1dminWindow_moreSmooth.rds')
```


## Plots - min window 7d

```{r}
# Utility function for plotting
set_ymx <- function(xx, ymax=2){ifelse(xx>ymax, ymax, xx)}
tmin = 125
```


1. Explain simulation

```{r}
sim_df %>%
    filter(time < 250 & time > tmin) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_line(data = sim_df, aes(x = time, y = incidence), color = 'deepskyblue')+
  geom_line(data = sim_df, aes(x = time, y = delayed), lty = 2, color = 'blue3')+
  #geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
  #geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ylab('incident infections\nobserved + unobserved')+
  ylim(c(0, 30000))+
  ggtitle('Hypothetical epidemic') -> inc

sim_df %>%
    filter(time < 250 & time > tmin) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_line(data = sim_df, aes(x = time, y = true_rt), color = 'deepskyblue') +
  geom_hline(aes(yintercept = 1), lty = 3)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste(R[t], ' values used to simulate epidemic'))) -> R0

cowplot::plot_grid(inc, R0, align = 'hv', nrow = 2)
ggsave('Hypothetical epidemic.png', width = 7, height = 4.5, dpi = 300)
```


```{r}
# test1 %>%
#   filter(time>=tmin) %>%
#   mutate_at(.vars = vars(starts_with('rt.')), set_ymx) %>%
#   ggplot() +
#   geom_hline(aes(yintercept = 1))+
#   geom_line(aes(x = time, y = true_rt))+
#   geom_ribbon(aes(x = time, ymin = rt.lower, ymax = rt.upper, fill = last_day), alpha = .5)+
#   geom_line(aes(x = time, y = rt.mean, color = last_day)) +
#   geom_vline(aes(xintercept = as.numeric(as.character(last_day)), color = last_day))+
#   scale_color_viridis_d(aesthetics = c('color', 'fill'))+
#   facet_wrap(.~median_cases_per_day)
```

```{r}
# tt_labs = c('Rt crosses 1\n(t=195)', '+1 week', '+2 weeks', '+3 weeks')
# names(tt_labs) = c(195, 202, 209, 216)

mm_labs = sprintf('%.0f', med_daily_obs)
names(mm_labs) = med_daily_obs

## Get a data frame of gaps between last available Rt estimate and actual last date
# test1 <- read_rds('power_results_7dminWindow.rds.rds')
# segment_df <- test1 %>%
#   group_by(median_cases_per_day, last_day) %>%
#   summarise(last_obs = unique(as.numeric(as.character(last_day))),
#             last_rt_est = max(time)) %>%
#   mutate(yy = .4+.05*(as.numeric(last_day)-2.5))
# 
# 
# test1 %>%
#   filter(time>=tmin) %>%
#   mutate_at(.vars = vars(starts_with('rt.')), set_ymx) %>%
#   pivot_longer(c(rt.mean, true_rt), names_to = 'kind', values_to = 'Rt') %>%
#   ggplot() +
#   geom_hline(aes(yintercept = 1), lty = 2)+
#   geom_ribbon(aes(x = time, ymin = rt.lower, ymax = rt.upper), alpha = .5, fill = 'dodgerblue')+
#   geom_line(aes(x = time, y = Rt, color = kind)) +
#   scale_color_manual('', values = c('dodgerblue', 'black'), aesthetics = c('color', 'fill'), labels = c('estimate', 'actual'))+
#   geom_line(data = sim_df, aes(x = time, y = true_rt))+
#   #geom_segment(data = segment_df, aes(x = last_rt_est, xend = last_obs, y = yy, yend = yy, color = last_day))+
#   facet_grid(.~median_cases_per_day, labeller = labeller(median_cases_per_day = mm_labs))+
#   xlim(c(tmin, 230))+
#   ylab('Rt')+
#   theme(legend.position = 'top')
# ggsave('Sens_results_7dminWindow.png', width = 7, height = 2.5, dpi = 300)


test2 <- read_rds('power_results_1dminWindow_moreSmooth.rds')

gapdf <- test2 %>%
  group_by(median_cases_per_day) %>%
  summarise(mt = max(time),
            gap = 250-mt)
test2 %>%
  filter(time>=tmin) %>%
  mutate_at(.vars = vars(starts_with('rt.')), set_ymx) %>%
  pivot_longer(c(rt.mean, true_rt), names_to = 'kind', values_to = 'Rt') %>%
  ggplot() +
  geom_hline(aes(yintercept = 1), lty = 2)+
  geom_ribbon(aes(x = time, ymin = rt.lower, ymax = rt.upper), alpha = .5, fill = 'dodgerblue')+
  geom_line(aes(x = time, y = Rt, color = kind)) +
  scale_color_manual('', values = c('dodgerblue', 'black'), aesthetics = c('color', 'fill'), labels = c('estimate', 'actual'))+
  geom_line(data = sim_df, aes(x = time, y = true_rt))+
  geom_vline(aes(xintercept = truncate.times), color = 'dodgerblue', lty = 2)+
  geom_label(data = gapdf, aes(x = 200, y = 2, label = sprintf('%.0fd gap', gap))) +
  #geom_segment(data = segment_df, aes(x = last_rt_est, xend = last_obs, y = yy, yend = yy, color = last_day))+
  facet_grid(.~median_cases_per_day, labeller = labeller(median_cases_per_day = mm_labs))+
  xlim(c(tmin, 250))+
  ylab('Rt')+
  theme(legend.position = 'top')
ggsave('Sens_results_1dminWindow.png', width = 7, height = 2.5, dpi = 300)
```


## Calculate mean delay 

```{r}
# test1 %>%
#   filter(time >= 125) %>%
#   group_by(median_cases_per_day, last_day) %>%
#   summarise(detection_lag = unique(as.numeric(as.character(last_day)))-max(time),
#             median_CI_width = median(rt.upper-rt.lower),
#             lower_CI_width = quantile(rt.upper-rt.lower, probs = .025),
#             upper_CI_width = quantile(rt.upper-rt.lower, probs = .8),
#             highest.lt1 = 1-upper_CI_width/2,
#             lowerst.ht1 = 1+upper_CI_width/2) -> stbl
# 
# stbl
# 
# 
# stbl %>%
#   ggplot()+
#   geom_rect(aes(xmin = 0, xmax = 1, ymin = highest.lt1, ymax = lowerst.ht1, fill = median_cases_per_day), alpha = .7)+
#   geom_segment((aes(x = 0, xend = 1, y = 1, yend = 1)))+
#   ylim(c(0, 2.5))+
#   scale_fill_viridis_d()
# 
# stbl %>%
#   ggplot(aes(x = as.numeric(as.character(median_cases_per_day)), y = upper_CI_width)) +
#   geom_point() +
#   geom_line(lty = 2) +
#   ylab('CI width (80th percentile)')+
#   xlab('Median cases/day')+
#   ylim(c(0, 1))
# ggsave('80th_pct_CI_7dminWindow.png', width = 2.5, height = 2.5, dpi = 300)
```

```{r}
test2 %>%
  filter(time >= 125) %>%
  group_by(median_cases_per_day, last_day) %>%
  summarise(detection_lag = unique(as.numeric(as.character(last_day)))-max(time),
            median_CI_width = median(rt.upper-rt.lower),
            lower_CI_width = quantile(rt.upper-rt.lower, probs = .025),
            upper_CI_width = quantile(rt.upper-rt.lower, probs = .8),
            highest.lt1 = 1-upper_CI_width/2,
            lowerst.ht1 = 1+upper_CI_width/2) -> stbl

stbl


stbl %>%
  ggplot()+
  geom_rect(aes(xmin = 0, xmax = 1, ymin = highest.lt1, ymax = lowerst.ht1, fill = median_cases_per_day), alpha = .7)+
  geom_segment((aes(x = 0, xend = 1, y = 1, yend = 1)))+
  ylim(c(0, 2.5))+
  scale_fill_viridis_d()

stbl %>%
  ggplot(aes(x = as.numeric(as.character(median_cases_per_day)), y = upper_CI_width)) +
  geom_point() +
  geom_line(lty = 2) +
  ylab('CI width (80th percentile)')+
  xlab('Median cases/day')+
  ylim(c(0, 1))
ggsave('80th_pct_CI_1dminWindow.png', width = 2.5, height = 2.5, dpi = 300)
```

