---
title: "Test sample size"
output: html_notebook
---

Katie Gostic

```{r asis = T}
cat(sprintf('\n## Last run %s', Sys.Date()))
```
```{r}
## Preamble
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
```



## Set observation delays
```{r}
rdelay <- function(nn){
  rlnorm(nn, meanlog = 1.52, sdlog = sqrt(0.39)) %>%  # Incubation period from table 2 of Linton et al.
  #runif(nn, 1, 2) %>% ## 1-4d reporting delay
  ceiling()
}

```


## Visualize synthetic data
```{r echo = F}
source('test_sample_size.R')
source('../code/util.R')
parlist <- read_rds('true_pars.rds')
sim_df <- get_obs_ts(rdelay)
sim_df %>%
    filter(time < 300) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_line(data = sim_df, aes(x = time, y = incidence), color = 'deepskyblue')+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Epidemic curve') -> inc

sim_df %>% 
  filter(time < 300) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_line(data = sim_df, aes(x = time, y = true_rt), color = 'deepskyblue') +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste('Underlying ', R[t], ' values'))) -> R0

cowplot::plot_grid(R0, inc, align = 'hv', nrow = 2)


filter(sim_df, time <200 & time > 175)
```

## Get the times at which Rt crosses one
```{r}
get_rt_cross_1_dates(sim_df)
rt.rise.time = get_rt_cross_1_dates(sim_df) %>% slice(2) %>% pull(time)
```
Debug - Test Cori method.
```{r}
# source('../code/cori.R')
# test_cori <-  merge(
#   get_sim_df('stochastic'), 
#     get_cori(df.in = get_sim_df('stochastic'),
#          icol_name = 'incidence', 
#          out_name = 'test', 
#          window = 1, 
#          GI_mean = parlist$true_mean_GI, 
#          GI_var = parlist$true_var_GI, 
#          wend = F),
#   by = 'time', 
#   all = T)
# 
# test_cori %>%
#   ggplot(aes(x = time))+
#   geom_line(aes(y = true_rt))+
#   geom_line(aes(y = test.mean), color = 'blue')
```



## Test various sample sizes

5000/d 
2000/d - total observed in IL as of 25 Sept.
500/d - total observed in Chicago as of 25 Sept
100/d
50/d
10/d


* Truncate data so the last time of observation falls at the time Rt first rises back above 1, 7, 14 and 21 days later.
* Estimate Rt using each truncated data set.
* Plot results


## Run ss pipeline with 2d min window
```{r error=FALSE, warning=FALSE}
truncate.times <- 250 # rt.rise.time + c(0, 7, 14, 21)
med_daily_obs <- c(4000, 2000, 500, 100, 50, 10)
test_list = expand.grid(tt = truncate.times, med_cases = med_daily_obs)
```

```{r error=FALSE, warning=FALSE}
wrapper1 <- function(tt, med_cases){test_sample_size_end_date(p_obs = 1,
                                                  median_cases_per_day = med_cases,
                                                  nboot = 100,
                                                  last_obs_time = tt,
                                                  rdelay = rdelay,
                                                  min_window = 2, 
                                                  w.tune = 50,
                                                  uncertainty = 'ss_pipeline'
                                                  )}
test1 <- mapply(FUN = wrapper1, tt = test_list$tt, med_cases = test_list$med_cases, SIMPLIFY = 'list') %>%
  apply(2, bind_cols) %>%
  bind_rows() %>%
  mutate(last_day = factor(last_day, levels = sort(truncate.times, decreasing = F)),
         median_cases_per_day = factor(median_cases_per_day, levels = sort(med_daily_obs, decreasing = F)))
write_rds(test1, path = 'results_sspipeline_2dminWindow.rds')
```

##  Run upscale pipeline with 2d min window
```{r error=FALSE, warning=FALSE}
wrapper2 <- function(tt, med_cases){test_sample_size_end_date(p_obs = 1,
                                                  median_cases_per_day = med_cases,
                                                  nboot = 100,
                                                  last_obs_time = tt,
                                                  rdelay = rdelay,
                                                  min_window = 1,
                                                  w.tune = 50,
                                                  uncertainty = 'upscale_pipeline'
                                                  )}
test2 <- mapply(FUN = wrapper2, tt = test_list$tt, med_cases = test_list$med_cases, SIMPLIFY = 'list') %>%
  apply(2, bind_cols) %>%
  bind_rows() %>%
  mutate(last_day = factor(last_day, levels = sort(truncate.times, decreasing = F)),
         median_cases_per_day = factor(median_cases_per_day, levels = sort(med_daily_obs, decreasing = F)))
write_rds(test2, path = 'results_upscalepipeline_2dminWindow.rds')
```

