---
title: "Estimate Rt from incident deaths"
output: html_notebook
---

Katie Gostic + Phil Arevalo

```{r asis = T}
cat(sprintf('\n## Last run %s', Sys.Date()))
```


```{r echo=FALSE}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE)
source('../code/util.R')
source('../code/cori.R')
```


## Load data
```{r}
dat <- read_csv('../data/idph_linelist_timeseries.csv') %>%
  filter(restore_region != 'unknown') %>%
  group_by(date, restore_region) %>%
  summarise(new_deaths = sum(new_deaths))%>%
  ungroup
dat %>%
  ggplot()+
  geom_line(aes(x = date, y = new_deaths))+
  facet_wrap(.~restore_region, scales = 'free_y')
```

# 1. Bootstrap the observed time series

** TO DO **

# 2. Sample across uncertainty in the delay distribution
```{r}
mean_delay <- 
var_mean <- 
var_delay <-
var_var <- 
```



## Function to estimate total infections given a fixed probability of observation
```{r}
est_total_inf <- function(df, ## Data frame containing a column with daily observations
                          p_obs, ## Vector of observation probabilities, one entry for each observation. If you provide a single number, assumes constant observation probability
                          obs_colname, ## Name of the column holding the observations
                          n_replicates = 1) ## Number of times to resample total infections 
  {
  draw_one_sample <- function(k, p){
    # adjust observations to work with negative binomial, i.e., case with 0 observations
    adjusted_k = replace(k, k==0, 0.1)
    
    # Draw a negative binomial sample
    observations = round(mapply(rnbinom, 1, adjusted_k, p) + adjusted_k)
    as.numeric(observations)
  }
  
  obs = df %>% select(obs_colname) %>% unlist(use.names = F)
  
  ## For each entry in the observations column, upscale
  infection_samples = replicate(n = n_replicates, draw_one_sample(obs, p_obs))

  colnames(infection_samples) = paste0('infections.', 1:n_replicates)
  bind_cols(df, as.data.frame(infection_samples))
}
```

##  Example -- draw one time series of infection

* red shows estimated infections
* blue shows observations

```{r}
## Get an aggregate time series of observations for the whole state
all_IL <- dat %>% group_by(date) %>% summarise(new_deaths = sum(new_deaths))
## Estimate total infections from observations
est_total_inf(all_IL, 
              p_obs = .5, 
              obs_colname = 'new_deaths',
              n_replicates = 1) %>%
  pivot_longer(-date) %>%
  ggplot()+
  geom_line(aes(x = date, y = value, color = name))+
  ggtitle('p = .5')
```

## Example -- bootstrap many time series of infections

* red shows estimated infections
* blue shows observations

```{r}
## Example -- bootstrap many time series of infections from observations
est_total_inf(all_IL, 
              p_obs = .1, 
              obs_colname = 'new_deaths', 
              n_replicates = 50) %>%
  pivot_longer(-date) %>%
  mutate(is.observed = !grepl('infection', name)) %>%
  ggplot() +
  geom_line(aes(x = date, y = value, group = name, color = is.observed), alpha = .5)+
  geom_hline(aes(yintercept = 1))
```
## Wrapper to feed resampled time series of observations into Rt estimator
```{r}
overall_rt_boot_wrapper <- function(
 dat,
 obs_colname,
 p_obs,
 mean_delay,
 n_replicates
){
  
  ## Rename the observations column for ease of manipulation
  dat <- rename(dat, observations = eval(obs_colname))
  
  ## Calculate the window size
  low_inf_count = filter(dat, observations>0) %>%
    pull(observations) %>%
    quantile(.2) %>%
    round()
  ww = floor(50/low_inf_count)
  cat(sprintf('\noverall window is %.0f\n', ww))
  
  ## Estimate total infections from observations
  infection_ests <- est_total_inf(df = dat, obs_colname = 'observations', p_obs = p_obs, n_replicates = n_replicates)
  

  
  ## Estimate Rt and mrege with the cleaned input data frame
  library(doParallel)
  cl <- makeCluster(parallel::detectCores()-1)
  registerDoParallel(cl)
  est_list <- foreach(inf_col= paste0('infections.', 1:n_replicates),
                      .packages = 'dplyr','EpiEstim', 'tidyr', 
                      .export = c('get_cori', 'na_to_0')) %dopar% {
                          ins = infection_ests %>% 
                            select(date, eval(inf_col)) %>%
                            mutate(time = 1:nrow(.))
                   
                            get_cori(ins, 
                              obs_col_name = inf_col, 
                              window = ww, 
                              out_name = 'rt',
                              mean_delay = mean_delay, 
                              SI_mean = 4.5, ## Rough estimates from Ganyani et al
                              SI_var = 1.7,   
                              wend = F) %>%
                            merge(select(ins, time, date), by = 'time') %>%
                            select(-time)
  }
  stopCluster(cl)
  
  ## For each date, get the median of means, the lower .025 of lower bounds and the upper .975 of upper bounds
  summarized_ests <-  bind_rows(est_list, .id = 'replicate') %>%
    group_by(date) %>%
    ## PHIL - for each replicate, we get a mean, upper and lower CI rt estimate.
    ##    I'm not sure what the best way to aggregate is.
    ##    Here I'm taking the median of means, the lower bound of lower bounds and the upper bound of upper bounds.
    summarise(rt.mean = median(rt.mean, na.rm = TRUE),
              rt.lower = quantile(rt.025, probs = .025, na.rm = TRUE),
              rt.upper = quantile(rt.975, probs = 0.975, na.rm = TRUE))
  
  return(summarized_ests)
  
}


```


## Test bootstrap
```{r}
test_overall_bootstrap <- overall_rt_boot_wrapper(dat = all_IL, 
                        obs_colname = 'new_deaths', 
                        p_obs = .8, 
                        mean_delay = 20, 
                        n_replicates = 50)
```

```{r}
test_overall_bootstrap %>%
  ggplot()+
  geom_line(aes(x = date, y = rt.mean))+
  geom_ribbon(aes(x = date, ymin = rt.lower, ymax = rt.upper), alpha = .3)+
  geom_hline(aes(yintercept = 1))+
  ylab('Rt')
```

