---
title: "Test RL on synthetic data using uncertain delay"
output: html_notebook
---

Katie Gostic + Phil Arevalo

```{r asis = T}
cat(sprintf('\n## Last run %s', Sys.Date()))
```
```{r}
## Preamble
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE)
parlist <- read_rds('../examples_syndat/true_pars.rds')
source('../code/util.R')
source('../code/cori.R')
source('../code/Richardson_Lucy.R')
source('../code/deconvolve.R')
source('../code/upscale.R')
source('../code/rt_boot.R')
if(!dir.exists('figs')){dir.create('figs')}
outdir = 'reorder'
```

## Extract SEIR simulation and plot.

- Incidence here is observed at time of infection (S -> E).
- SEIR simulation is deterministic

```{r}
setwd('../examples_syndat/')
get_sim_df() %>%
    filter(time < 300) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Epidemic curve') -> inc

get_sim_df() %>% 
  filter(time < 300) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste('Underlying ', R[t], ' values'))) -> R0

cowplot::plot_grid(R0, inc, align = 'hv', nrow = 2)

sim_df <- get_sim_df()
```


# Fig. 2 - estimate Rt in real time using three different methods, assuming perfect observation at the moment of infection.

## Truncate the synthetic data to a maximum time of observation

```{r}
# 
# post <- read_rds('../data/fitted_delays/delay_infection_to_death_posterior.rds') %>%
#   bind_cols() %>%
#   arrange(mu)
# 
# icolname <- 'incidence'
# delay_type = 'lognormal'
# true_delay_pars <- c(mu = post$mu[round(nrow(post)/2)],
#                      sigma = post$sigma[round(nrow(post)/2)])
# delay_posterior <- post
# mt = 150
# p_obs = .5


full_test <- function(sim_df,
                      icolname,
                      mt, 
                      p_obs, 
                      true_delay_pars, 
                      delay_posterior, 
                      delay_type = 'lognormal',
                      gen_int_pars,
                      nboot = 100,
                      outname){
  
  ## Check inputs
  sim_df$incidence <- sim_df[,icolname]
  sim_df$incidence <- round(sim_df$incidence)
  sim_df$incidence <- na_to_0(sim_df$incidence)
  stopifnot('incidence' %in% colnames(sim_df))
  
  source('../code/infer_times_of_infection_observation.R')
  
  ## 1. Generate synthetic observations ----------
  ##     Set the delay to observation
  if(delay_type == 'lognormal'){
    rdelay <- function(nn){rlnorm(nn, true_delay_pars[1], true_delay_pars[2])}
  }else{
    stopifnot(delay_type == 'gamma')
     rdelay <- function(nn){rgamma(nn, true_delay_pars[1], true_delay_pars[2])}
  }
  
  ## Convolve to get synthetic times of observation
  sim_df <-  merge(
    sim_df %>% select(time, incidence, true_rt),
    get_tObs_from_tInf(n_dS = sim_df$incidence, 
                     times = sim_df$time, 
                     r_delay_dist = rdelay, 
                     return_times = T) %>% rename(delayed = n),
  by = 'time') %>%
  filter(time <= mt)
  ## Downsample based on pobs
  sim_df$obs = sapply(sim_df$delayed, FUN = function(NN) rbinom(n=1, size = NN, prob = p_obs))
  ## Shift back by mean delay
  sim_df$shifted_obs = lead(sim_df$obs, n = round(mean(rdelay(10000))))
  ## Plot
  sim_df %>%
    select(time, incidence, delayed, obs, shifted_obs) %>%
    setNames(c('time', '1. total infections', '2. total delayed', '3. observations', '4. shifted observations')) %>%
    pivot_longer(-time) %>%
    ggplot() +
    geom_line(aes(x = time, y = value, color = name))+
    ylab('daily count')+
    ggtitle('synthetic data') -> syndat
  
  ## 2. Bootstrap the observed time series ---------------
  
  ## 4. Upscale partial shifted observations
  upscaled <-   est_total_inf(sim_df, 
                p_obs = p_obs, 
                obs_colname = 'shifted_obs',
                n_replicates = nboot) 
  
  
  ## 3. Use RL to infer times of infection given the posterior parms of the delay
  deconvolved <- sapply(grep('infection', colnames(upscaled)), FUN = function(cc){
    
    ts = upscaled[,cc]
  
                        deconvolve(ts, 
                                   delay_posterior = delay_posterior, 
                                   nboot = 1, 
                                   delay_type = delay_type) 
  }
  )
  

  
  RL_plot <- syndat + 
    geom_line(data = deconvolved %>% bind_rows(.id = 'rep'), aes(x = time, y = RL_result, group = rep), alpha = .1) +
    ylim(c(0, max(sim_df$incidence)+20))+
    ggtitle('Inferred infections (partial)')
  
  
  upscale_plot <- syndat + 
    geom_line(data = upscaled %>% bind_rows(.id = 'rep'), aes(x = time, y = infections.1, group = rep), alpha = .1)+
        ylim(c(0, max(sim_df$incidence)+20))+
    ggtitle('Estimated total infections')
  
  
  ## Calculate the window size
  low_inf_count = filter(bind_rows(deconvolved), RL_result>0 & !is.na(RL_result)) %>%
    pull(RL_result) %>%
    quantile(.2) %>%
    round()
  ww.in = max(1, floor(100/low_inf_count))
  cat(sprintf('\noverall window is %.0f\n', ww.in))
  
  ## 5. Estimate Rt
  rt_ests <- rt_boot(infection_ests = bind_rows(deconvolved, .id = 'rep') %>% 
                       mutate(RL_result = round(RL_result)) %>%
                       pivot_wider(id_cols = time, 
                                   names_from = rep, 
                                   names_prefix = 'infections.', 
                                   values_from = RL_result),
                        p_obs = p_obs, ww = ww.in)
  
  
  ## 6. Format outputs
 rt_plot <- sim_df %>%
    merge(rt_ests$summary, by = 'time') %>% 
    ggplot()+
    geom_line(aes(x = time, y = true_rt))+
    geom_ribbon(aes(x = time, ymin = rt.lower, ymax = rt.upper), fill = 'yellow', alpha = .5)+
       geom_line(data = bind_rows(rt_ests$all, .id = 'rep'), aes(x = time, y = rt.mean, group = rep), alpha = .5, color = 'orange')+
   geom_line(aes(x = time, y = rt.mean), color = 'red')+
   ylim(c(0, 3))+
   ylab('Rt')
 
 
 delplot <- delay_posterior %>% 
   select(1:2) %>%
   pivot_longer(1:2) %>%
   ggplot()+
   geom_vline(data = data_frame(name = c('mu', 'sigma'), value = true_delay_pars), aes(xintercept = value))+
   geom_density(aes(x = value, fill = name, color = name), alpha = .5)+
   facet_grid(.~name, scales = 'free_x') +
   ggtitle(sprintf('%s delay', delay_type), subtitle = 'speficied distribution (colored distribution) vs. truth (black line)')
 
 
 cowplot::plot_grid(syndat, 
                    plot_grid(RL_plot+theme(legend.position = 'none'), 
                              upscale_plot+theme(legend.position = 'none'), nrow = 1), 
                    delplot,
                    rt_plot, nrow = 4, ncol = 1)
 ggsave(sprintf('testfigs/%s.png', 
                outname),
          width = 6, height = 8, dpi = 300, units = 'in')
 
 return(list(sim_df = sim_df %>% merge(rt_ests$summary, by = 'time'),
             RL = deconvolved,
             upcaled = upscaled,
             rt_ests = rt_ests))
  
}
```

## Set distribution of delay params
```{r}
## From Lauer et al
del_post = data.frame(mu = rnorm(1000, 1.621, .067), 
                      sigma = rnorm(1000, .418, 0.062)) %>%
  mutate(mu = ifelse(mu <= .001, .001, mu),
         sigma = ifelse(sigma <= 0.001, .001, sigma))

del_pars <- colMeans(del_post)
```


## Test realistic lognormal delay
Don't consider variability in the delay parameters
```{r}
# icolname = 'incidence'
# mt = 150
# p_obs = 1
# true_delay_pars = c(1.621, .418)
# delay_posterior = del_post
# delay_type = 'lognormal'
# gen_int_pars = c(parlist$true_mean_SI, parlist$true_var_SI)
# nboot = 10
# outname = 'test'

```

## partial observation
With partial observation
```{r}
test_3 <- full_test(sim_df, 
          icolname = 'incidence', 
          mt = 150, 
          p_obs = .5, 
          true_delay_pars = del_pars, 
          delay_posterior = del_post, 
          delay_type = 'lognormal', 
          gen_int_pars = c(parlist$true_mean_SI, parlist$true_var_SI), 
          nboot = 100,
          outname = paste0(outdir, '/perfectSpec_.5Obs_postVar'))
```


```{r}
test_4 <- full_test(sim_df, 
          icolname = 'incidence', 
          mt = 150, 
          p_obs = .1, 
          true_delay_pars = del_pars, 
          delay_posterior = del_post, 
          delay_type = 'lognormal', 
          gen_int_pars = c(parlist$true_mean_SI, parlist$true_var_SI), 
          nboot = 100,
          outname = paste0(outdir, '/perfectSpec_.1obs_postVar'))
```




```{r}
test_5 <- full_test(sim_df, 
          icolname = 'incidence', 
          mt = 150, 
          p_obs = .5, 
          true_delay_pars = c(log(exp(del_pars[1])+2), log(exp(del_pars[2])+3)), 
          delay_posterior = del_post, 
          delay_type = 'lognormal', 
          gen_int_pars = c(parlist$true_mean_SI, parlist$true_var_SI), 
          nboot = 100,
          outname = paste0(outdir, '/misspec2d_.5Obs_postVar'))
```

## Test realistic lognormal delay
With partial observation
```{r}
test_6 <- full_test(sim_df, 
          icolname = 'incidence', 
          mt = 150, 
          p_obs = .1, 
          true_delay_pars = c(log(exp(del_pars[1])+2), log(exp(del_pars[2])+3)), 
          delay_posterior = del_post, 
          delay_type = 'lognormal', 
          gen_int_pars = c(parlist$true_mean_SI, parlist$true_var_SI), 
          nboot = 100,
          outname = paste0(outdir, '/misspec2d_.1Obs_postVar'))
```