---
title: "Analyze synthetic data and generate plots"
output: html_notebook
---


```{r}
## Preamble
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE)
parlist <- read_rds('true_pars.rds')
source('../code/util.R')
source('../code/cori.R')
if(!dir.exists('figs')){dir.create('figs')}
```

## Extract SEIR simulation and plot.

- Incidence here is observed at time of infection (S -> E).
- SEIR simulation is deterministic

```{r}
get_sim_df() %>%
    filter(time < 300) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Epidemic curve') -> inc

get_sim_df() %>% 
  filter(time < 300) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste('Underlying ', R[t], ' values'))) -> R0

cowplot::plot_grid(R0, inc, align = 'hv', nrow = 2)
```


# Fig. 2 - estimate Rt in real time using three different methods, assuming perfect observation at the moment of infection.

## Truncate the synthetic data to a maximum time of observation

```{r}
mt <- 150
rtdf <- get_sim_df() %>% filter(time <= mt)
```


## Use the method of Cori et al. and of Wallinga & Teunis from EpiEstim to estimate Rt.
```{r}
wdw <- 1 ## Set smoothing window = 1d

plot_df_rt <- rtdf %>%
  merge(get_cori(rtdf, 
                 icol_name = 'incidence', 
                 window = wdw), 
        by = 'time') 
```

## Plot Figure 2
```{r}
plot_df_rt %>%
  ggplot() +
  geom_line(aes(x = time, y = true_rt), lwd = 1)+
  geom_line(aes(x = time, y = Cori.mean), color = 'dodgerblue', lty = 2) +
  geom_ribbon(aes(x = time, ymin = Cori.025, ymax = Cori.975), fill= 'dodgerblue',  alpha = 0.3)+
  xlab('time') +
  ylab(expression(paste(R[t]))) 
```

